Stochastic Variance Reduced Policy Gradient

### 2018-02-01 svrpg.tex
started writing intro

\section{Introduction}
\begin{itemize}
        \item Similarities between RL and SL
        \item SL needs SVRG so RL
        \item Potential advantages of SVRG
        \item Difficulties in RL \wrt SL
\end{itemize}

@@

Policy gradient methods restrict the search for the performance-maximizing policy over a class of parametrized policies $\Pi_{\vtheta}=\{\pol: \vtheta \in \realspace^m\}$, with the only constraint that $\pol$ is differentiable \wrt the $m$-dimensional parameter vector $\vtheta$. We denote the performance of a parametric policy with $J(\vtheta)$ for brevity, while in some occasions we replace $\pol(\tau)$ with $\pi(\tau|\vtheta)$ for the sake of readability. The search for a locally optimal policy is performed by means of gradient ascent, where the gradient of $J(\vtheta)$ \wrt the policy parameters is \cite{sutton2000policy}:
\todopir{stochastic or mini-batch gradient}
\begin{align*}
\gradJ{\vtheta} = \int_{\Tspace}\pol(\tau)\score{\vtheta}{\tau}\Reward(\tau)\de \tau.
\end{align*}
We consider the on-line learning problem, where trajectories must be sampled directly by interacting with the environment. 
In this setting, $(N>0)$-batch gradient ascent is typically employed: after sampling a batch of $N$ trajectories (\ie $\Dataset_{N}$),
policy parameters are updated as $\vtheta\gets\vtheta + \alpha\gradApp{\vtheta}{N}$, where $\alpha$ is a step size and $\gradApp{\vtheta}{N}$ is an estimate of the true gradient. 
\todopir{step size is scalar or vectorial (just for notation)? Check consistency of $N$ with finite-sum function, we can use even $b$ or $B$ for the batch size.}
The REINFORCE gradient estimator~\citep{williams1992simple} provides a simple, unbiased way of estimating the gradient:
\begin{align*}
\gradApp{\vtheta}{N} = \frac{1}{N}\sum_{n=1}^{N}\left(\sum_{h=0}^{H-1}\score{\vtheta}{a_h^n\mid s_h^n}\right)\left(\sum_{h=0}^{H-1}\gamma^h r_h^n - b\right),
\end{align*}
where subscripts denote the time step, superscripts denote the trajectory, $r_h$ is the reward actually collected at time $h$ and b can be any baseline, provided it is constant \wrt to actions.
\todopir{Actually it can be state-action dependant~\citep[\eg][]{Thomas2017actionbaseline,anonymous2018variance}.}
The G(PO)MDP gradient estimator~\cite{baxter2001infinite} is a refinement of REINFORCE which is subject to less variance while preserving the unbiasedness:
\begin{align*}
\gradApp{\vtheta}{N} = \frac{1}{N}\sum_{n=1}^{N}\sum_{h=0}^{H-1}\left(\sum_{k=0}^{h}\score{\vtheta}{a_k^n\mid s_k^n}\right)\left(\gamma^h r_h^n - b\right).
\end{align*}
\todopir{Cite Jan's paper somewhere}
\todopir{At this point we need to connect with the supervised learning framework (we can just define everything inside the first sum as $\nabla g(\tau)$).}

@@

\section{Hyperparameters}\label{sec:stopping}
In Reinforcement learning we have to take into account the gradient's variance.
A used way for reduce the effect of variance is to use Adaptive Moment Estimation (Adam) \cite{kingma2014adam}, a method that computes adaptive learning rates for each parameter. Adam help to stabilize the gradient estimate. In fact, this method computes adaptive learning rates for each parameter that depends on the history of the gradients' variance.
We use two different adam hyperparameters: one for updates tied to Snapshots points and the other tied to the updates of sub-iterations points. 
This is because we're dealing with two different variance dimensions.
Denote by $\alpha_{snap}$ the last learning rate computed into the snapshot point and denote by  $\alpha_{sub-iter}$ the last learning rate computed into the sub-iteration point.

As shown, the performance of the SVRPG algorithm is sensitive to the number of sub-iterations.
As the number of sub-iterations increases, the variance of the importance weights and the gradient increases too.
Therefore, the SVRPG algorithm fails in its primary aim of stabilizing the gradient estimate.
A fixed number of sub-iterations is not the optimal choice.
Few sub-iterations lead to few updates, many update lead to increases of variance.

We introduce two stopping condition: the maximum number of sub-iterations and the condition on learning rate of Adam sub-iterations.
We stop with sub-iterations if:

>>>>

\section{Introduction}
%Similarities between RL and SL
On a very general level, machine learning addresses the problem of artificial agents that must select the right actions to solve a given task. The approach of Reinforcement Learning \cite{sutton2011reinforcement} is to learn the best actions by direct interaction with the environment and evaluation of the performance in the form of a reward signal. This makes RL fundamentally different from Supervised Learning (SL), where correct actions are explicitly prescribed by a human teacher (\eg for classification, in the form of class labels). However, the two approaches share many challenges and, as a consequence, solution concepts. The problem of estimating a model from samples, which is at the core of SL, is equally fundamental in RL, whether we choose to model the environment, a value function, or a policy of behavior directly. In particular, when the tasks are characterized by large or continuous state-action spaces, RL needs to borrow the powerful function approximators that are the main subject of study of SL, such as neural networks. Among RL solutions for learning continuous tasks, policy gradient methods bear the closest similarity to SL, since they allow to formulate the RL problem as the unconstrained optimization of an objective function (\ie a negative loss) of policy parameters and data samples.

% SL needs SVRG so RL
In supervised learning, stochastic optimization is often employed for reasons of efficiency. The idea is to evaluate the objective on a random subset of the available data to obtain an approximate gradient direction. This can be much less expensive that taking into account all the data for all parameter updates. However, the sub-sampling of data samples is a source of variance that can potentially compromise convergence, so that speed and variance must be traded off with proper handling of meta-parameters.
Variance-reduced gradient algorithms such as SAG, SVRG and SAGA offer better ways of achieving the optimal trade-off between variance and speed, both in theory and practice. 
In Reinforcement Learning, stochastic optimization is not even a choice, since samples must be sequentially obtained by interacting with an initially unknown environment. In this scenario, containing the variance of the estimates is a problem that cannot be avoided, and methods like SVRG can offer a more efficient alternative

\todomat{work in progress}
% Potential advantages of SVRG
% Difficulties in RL \wrt SL

@@

Policy gradient methods restrict the search for the performance-maximizing policy over a class of parametrized policies $\Pi_{\vtheta}=\{\pol: \vtheta \in \realspace^m\}$, with the only constraint that $\pol$ is differentiable \wrt the $m$-dimensional parameter vector $\vtheta$. We denote the performance of a parametric policy with $J(\vtheta)$ for brevity, while in some occasions we replace $\pol(\tau)$ with $\pi(\tau|\vtheta)$ for the sake of readability. The search for a locally optimal policy is performed by means of gradient ascent, where the gradient of $J(\vtheta)$ \wrt the policy parameters is \cite{sutton2000policy}:
\todopir{stochastic or mini-batch gradient}
\begin{align*}
\gradJ{\vtheta} = \int_{\Tspace}\pol(\tau)\score{\vtheta}{\tau}\Reward(\tau)\de \tau.
\end{align*}
We consider the on-line learning problem, where trajectories must be sampled directly by interacting with the environment. 
In this setting, $(N>0)$-batch gradient ascent is typically employed: after sampling a batch of $N$ trajectories (\ie $\Dataset_{N}$),
policy parameters are updated as $\vtheta\gets\vtheta + \alpha\gradApp{\vtheta}{N}$, where $\alpha$ is a step size and $\gradApp{\vtheta}{N}$ is an estimate of the true gradient. 
\todopir{step size is scalar or vectorial (just for notation)? Check consistency of $N$ with finite-sum function, we can use even $b$ or $B$ for the batch size.}
The REINFORCE gradient estimator~\citep{williams1992simple} provides a simple, unbiased way of estimating the gradient:
\begin{align*}
\gradApp{\vtheta}{N} = \frac{1}{N}\sum_{n=1}^{N}\left(\sum_{h=0}^{H-1}\score{\vtheta}{a_h^n\mid s_h^n}\right)\left(\sum_{h=0}^{H-1}\gamma^h r_h^n - b\right),
\end{align*}
where subscripts denote the time step, superscripts denote the trajectory, $r_h$ is the reward actually collected at time $h$ and b can be any baseline, provided it is constant \wrt to actions.
\todopir{Actually it can be state-action dependant~\citep[\eg][]{Thomas2017actionbaseline,wu2018variance}.}
The G(PO)MDP gradient estimator~\cite{baxter2001infinite} is a refinement of REINFORCE which is subject to less variance while preserving the unbiasedness:
\begin{align*}
\gradApp{\vtheta}{N} = \frac{1}{N}\sum_{n=1}^{N}\sum_{h=0}^{H-1}\left(\sum_{k=0}^{h}\score{\vtheta}{a_k^n\mid s_k^n}\right)\left(\gamma^h r_h^n - b\right).
\end{align*}
\todopir{Cite Jan's paper somewhere}
\todopir{At this point we need to connect with the supervised learning framework (we can just define everything inside the first sum as $\nabla g(\tau)$).}

@@

\section{Stopping Condition}\label{sec:stopping}
As shown, the performance of the SVRPG algorithm is sensitive to the number of sub-iterations.
As the number of sub-iterations increases, the variance of the importance weights and the gradient increases too.
Therefore, the SVRPG algorithm fails in its primary aim of stabilizing the gradient estimate.
A fixed number of sub-iterations is not the optimal choice.
Few sub-iterations lead to few updates, many update lead to increases of variance.

We introduce two stopping condition: the maximum number of sub-iterations and the condition on learning rate of Adam sub-iterations.
Adam update rule help to stabilize the gradient estimate. In fact, this method computes adaptive learning rates for each parameter that depends on the history of the gradients' variance.
In particular we use two different Adam parameters: the first one for snapshot updates and the second one for the sub-iterations updates.
We stop with sub-iterations if:
